---
output:
  github_document:
    html_preview: false
always_allow_html: true
---

```{r,echo=FALSE}
knitr::opts_chunk$set(
  collapse=TRUE,
  comment="#>",
  fig.path="man/figures/"
)
```

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- Use snippet 'render_markdown' for it -->

# vegtable

<!-- badges: start -->
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/vegtable)](https://cran.r-project.org/package=vegtable)
[![Rdoc](http://www.rdocumentation.org/badges/version/vegtable)](http://www.rdocumentation.org/packages/vegtable)
[![DOI](https://zenodo.org/badge/55006983.svg)](https://zenodo.org/badge/latestdoi/55006983)
<br>
[![R-CMD-check](https://github.com/kamapu/vegtable/workflows/R-CMD-check/badge.svg)](https://github.com/kamapu/vegtable/actions)
[![codecov](https://codecov.io/gh/kamapu/vegtable/branch/master/graph/badge.svg)](https://codecov.io/gh/kamapu/vegtable)
<br>
[![CRAN_downloads](http://cranlogs.r-pkg.org/badges/vegtable)](https://cran.r-project.org/package=vegtable)
[![total downloads](http://cranlogs.r-pkg.org/badges/grand-total/vegtable)](https://cran.r-project.org/package=vegtable)
<!-- badges: end -->

The aim of `vegtable` is to provide a way for handling databases stored in
[Turboveg](http://www.synbiosys.alterra.nl/turboveg). This package incorporates
many concepts and some functions included in the package
[vegdata](https://cran.r-project.org/package=vegdata) but defining an homonymous
`S4` class containing all elements of a database in just one object. The package
`vegtable` also contains several methods for this object class.

Species lists in `vegtable` objects are handled by the package
[taxlist](https://github.com/kamapu/taxlist), thus I will recommend to take a
look on it.

This package has been developed as a tool handling data stored in
[SWEA-Dataveg](http://www.givd.info/ID/AF-00-006). Further development is
running in the context of the project
[GlobE-wetlands](https://www.wetlands-africa.de/).

An important source of inspiration for `vegtable` have been the enthusiastic
discussions during several versions of the [Meetings on Vegetation
Databases](http://www.hswt.de/person/joerg-ewald/vegetationsdatenbanken.html).

## Updating to the last version of vegtable
The very first step is to install the package
[devtools](https://github.com/hadley/devtools) and dependencies. Then you just
need to execute following commands in your R-session:

```{r, eval=FALSE}
library(devtools)
install_github("kamapu/vegtable")
```

## Some examples
The current version of `vegtable` includes an example data, which corresponds to
a subset from [SWEA-Dataveg](http://www.givd.info/ID/AF-00-006). This data set
contains plot observations done in Kenya imported from 5 sources.

```{r}
library(vegtable)
data(Kenya_veg)

# validate and explore
validObject(Kenya_veg)
summary(Kenya_veg)
```

Among others, the object contains plot observations done in the Aberdare
National Park (Kenya) by __Schmitt (1991)__. We can make a subset including the
plots classified by the mentioned author into the *Juniperus
procera*-*Podocarpus latifolius* community (IDs 780 to 798).

```{r}
JPcomm <- subset(Kenya_veg, ReleveID %in% c(780:798))
summary(JPcomm)
```

If you have geo-referenced plot observations, you can use the coordinates to
produce a map of the distribution of your plots by using the package `leaflet`.

```{r}
library(leaflet)
leaflet(JPcomm@header) %>%
		addTiles() %>%
		addCircleMarkers(lng = ~LONGITUDE, lat = ~LATITUDE, color = "red",
				opacity = 0.3, radius = 1)
```

## Further readings

- [Basics on the work with vegetation-plots in
  vegtable](https://kamapu.github.io/posts/2020-11-20-vegtablepress2/)
- [Introduction to the package
  vegtable](https://kamapu.github.io/posts/2020-10-29-vegtableintro/)
- [vegtable: An R object for vegetation-plot data
  sets](https://kamapu.github.io/posts/2020-10-29-vegtablepress/)
